# Question 1: What was the frequency of the following key words per year and the whole period?

„klima“, „globální oteplování“, „skleníkový efekt“, „uhlíková stopa“.

## Load necessary libraries and specify paths
```{r}
packages <-
  c(
    "dplyr",
    "stringr",
    "purrr",
    "tidyr",
    "parallel",
    "lubridate",
    "ggplot2"
  )

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, library, character.only = TRUE))

doc_id_by_date <- readRDS("data/doc_id_by_date.rds")

list_of_processed_chunks <- list.files(path = "../2.data_transformations/data/udpipe_processed", pattern = "*.rds", full.names = TRUE)

```

## Specify term counting function
```{r}
count_terms_of_interest <- function(text_vector, term) {
  str_count(text_vector, regex(term))
}

get_term_counts <- function(chunk_path) {
  text_stats <- chunk_path %>%
    readRDS() %>%
    filter(upos %in% c("ADJ", "NOUN", "PUNCT")) %>%
    transmute(doc_id, lemma = tolower(lemma)) %>%
    inner_join(doc_id_by_date, by = "doc_id") %>%
    mutate(date = as.character(year(date))) %>%
    group_by(doc_id) %>%
    summarize(
      text = str_squish(str_c(lemma, collapse = " ")),
      year = first(date),
      climate_count = count_terms_of_interest(text, "(?<!společenské |politické )klima\\b"),
      warming_count = count_terms_of_interest(text, "\\boteplování"),
      global_warming_count = count_terms_of_interest(text, "\\bglobální oteplování"),
      greenhouse_effect_count = count_terms_of_interest(text, "\\bskleníkový efekt"),
      carbon_footprint_count = count_terms_of_interest(text, "\\buhlíkový stopa"),
      ecological_count = count_terms_of_interest(text, "\\bekologický"),
      climatic_count = count_terms_of_interest(text, "\\bklimatický"),
      coal_count = count_terms_of_interest(text, "\\buhlí"),
      weather_count = count_terms_of_interest(text, "\\bpočasí"),
      combined_count = climate_count + warming_count + global_warming_count + greenhouse_effect_count + carbon_footprint_count + ecological_count + climatic_count + coal_count + weather_count
    ) %>%
    ungroup() %>%
    filter(combined_count > 0) %>%
    select(-c("text"))
}
```

## Run the function on chunk-to-chunk basis
```{r}

# If we are using Linux or MacOS, we can use multiple CPUs to make the whole
# process much faster. Each core handles different chunk at the same time.

if (Sys.info()[['sysname']] == "Windows") {
  # Use normal apply with Windows
  term_counts_df <-
    lapply(list_of_processed_chunks, get_term_counts) %>% bind_rows()

} else {
  term_counts_df <-
    mclapply(list_of_processed_chunks, get_term_counts, mc.cores = detectCores() - 2) %>% bind_rows()
}

saveRDS(term_counts_df, "data/term_counts_df.rds")
```

## Summarize the key term count per year
```{r}
term_counts_per_year <- term_counts_df %>%
  group_by(year) %>%
  summarise(
    n_climate = sum(climate_count),
    n_warming = sum(warming_count),
    n_global_warming = sum(global_warming_count),
    n_greenhouse_effect = sum(greenhouse_effect_count),
    n_carbon_footprint = sum(carbon_footprint_count),
    n_ecological = sum(ecological_count),
    n_climatic = sum(climatic_count),
    n_coal = sum(coal_count),
    n_weather = sum(weather_count)
  ) %>%
  ungroup() %>%
  bind_rows(summarise(
    ., across(where(is.numeric), sum),
    across(where(is.character), ~"Total")
  ))

knitr::kable(term_counts_per_year, format = "pipe")
```

## Visualize all of the terms and their frequencies
```{r}
term_counts_per_year %>%
  filter(year != "Total") %>% 
  pivot_longer(where(is.numeric), values_to = "count", names_to = "term") %>% 
  mutate(year = as.factor(year),
         term = str_remove(term, "^n_")) %>% 
  ggplot(aes(x = year, y = count, fill = term)) +
  geom_col() +
   theme_minimal() +
    scale_y_continuous(
      breaks = seq(0, 10000, 500),
      labels = seq(0, 10000, 500)
    ) +
    scale_fill_brewer(palette = "Set1")+ 
    labs(x = element_blank(),
             y = "Lemma occurences per year",
             fill = "")
```

## Visualize Just the main 4 terms
```{r}
term_counts_per_year %>%
  filter(year != "Total") %>% 
  pivot_longer(where(is.numeric), values_to = "count", names_to = "term") %>% 
  mutate(year = as.factor(year),
         term = str_remove(term, "^n_")) %>% 
  filter(term %in% c("global_warming", "climate", "greenhouse_effect", "carbon_footprint")) %>% 
  ggplot(aes(x = year, y = count, fill = term)) +
  geom_col() +
   theme_minimal() +
    scale_y_continuous(
      breaks = seq(0, 10000, 100),
      labels = seq(0, 10000, 100)
    ) +
    scale_fill_brewer(palette = "Set1")+ 
    labs(x = element_blank(),
             y = "Lemma occurences per year",
             fill = "")
```
