---
title: "More_term_frequencies"
author: "Irene Elmerot / Ondra Pekáček"
date: "2023-10-09"3
output: html_document
---

# This R Markdown file contains code for analyzing the frequency of key words related to climate change,
# including economic terms alongside "klima" and "klimatický". 
# The code loads necessary libraries, specifies paths, defines term counting functions, runs the function on a chunk-to-chunk basis, 
# summarizes the key term count per year, and visualizes the terms and their frequencies. 
# The code also includes a visualization of the main terms.
# The output of the code includes a table of the term counts per year and two plots showing the frequency of the key terms over time.

# Question 2: What was the frequency of the following key words per year and the whole period?

"Banka", "Koruna", "Burza", "Akcie", "Úvěr", "Exekuce", "Chudoba", "klima", "klimatický".

## Load necessary libraries and specify paths
```{r}
packages <-
  c(
    "dplyr",
    "stringr",
    "purrr",
    "tidyr",
    "parallel",
    "lubridate",
    "ggplot2"
  )

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, library, character.only = TRUE))

doc_id_by_date <- readRDS("data/doc_id_by_date.rds")

list_of_processed_chunks <- list.files(path = "../2.data_transformations/data/udpipe_processed", pattern = "*.rds", full.names = TRUE)

```

## Specify term counting function
```{r}
# This function counts the frequency of a given term in a vector of text.
# Args:
#   text_vector: A vector of text to search for the given term.
#   economic_term: A string representing the term to search for.
# Returns:
#   An integer representing the frequency of the given term in the text vector.
# Example:
#   count_terms_of_interest(c("The quick brown fox", "jumps over the lazy dog"), "the")
#   # Output: 2

count_economy_terms_of_interest <- function(text_vector, economy_term) {
  str_count(text_vector, regex(economy_term))
}

# This function reads in a chunk of data from a specified path and performs text analysis on it.
# It filters the text by parts of speech (ADJ, NOUN, PUNCT), converts all lemmas to lowercase, and joins the data with a document ID by date. 
# It then calculates the frequency of various climate-related terms in the text, 
# including "klima", "oteplování", "globální oteplování", "skleníkový efekt", "uhlíkový stopa", "ekologický", "klimatický", "uhlí", and "počasí". 
# The function returns a data frame with the document ID, year, and the frequency of each term in the text. NB: For bigrams, make sure both words are lemmatized, e.g. "český koruna" or "klimatický změna".
get_economy_term_counts <- function(chunk_path) {
  text_stats <- chunk_path %>%
    readRDS() %>%
    filter(upos %in% c("ADJ", "NOUN", "PUNCT")) %>%
    transmute(doc_id, lemma = tolower(lemma)) %>%
    inner_join(doc_id_by_date, by = "doc_id") %>%
    mutate(date = as.character(year(date))) %>%
    group_by(doc_id) %>%
    summarize(
      text = str_squish(str_c(lemma, collapse = " ")),
      year = first(date),
      climate_count = count_economy_terms_of_interest(text, "(?<!společenské |politické )klima\\b"),
      climatic_count = count_economy_terms_of_interest(text, "\\bklimatický"),
      bank_count = count_economy_terms_of_interest(text, "\\bbanka"),
      koruna_count = count_economy_terms_of_interest(text, "\\bčeský koruna"),
      exchange_count = count_economy_terms_of_interest(text, "\\bburza"),
      stockshare_count = count_economy_terms_of_interest(text, "\\bakcie"),
      credit_loan_count = count_economy_terms_of_interest(text, "\\búvěr"),
      foreclosure_count = count_economy_terms_of_interest(text, "\\bexekuce"),
      poverty_count = count_economy_terms_of_interest(text, "\\bchudoba"),
      combined_count = climate_count + climatic_count + bank_count + koruna_count + exchange_count + stockshare_count + credit_loan_count + foreclosure_count + poverty_count
    ) %>%
    ungroup() %>%
    filter(combined_count > 0) %>%
    select(-c("text"))
}
```

## Run the function on chunk-to-chunk basis
```{r}

# If we are using Linux or MacOS, we can use multiple CPUs to make the whole
# process much faster. Each core handles different chunk at the same time.

if (Sys.info()[['sysname']] == "Windows") {
  # Use normal apply with Windows
  economy_term_counts_df <-
    lapply(list_of_processed_chunks, get_economy_term_counts) %>% bind_rows()

} else {
  economy_term_counts_df <-
    mclapply(list_of_processed_chunks, get_economy_term_counts, mc.cores = detectCores() - 2) %>% bind_rows()
}

saveRDS(economy_term_counts_df, "data/economy_term_counts_df.rds")
```

## Summarize the key term count per year
```{r}
# This code chunk groups the term counts by year and summarizes the counts for each term.
# The resulting data frame has one row for each year, with columns for each term and the count of that term for that year.
economy_term_counts_per_year <- economy_term_counts_df %>%
  group_by(year) %>%
  summarise(
    n_climate_noun = sum(climate_count),
    n_climate_adj = sum(climatic_count),
    n_bank = sum(bank_count),
    n_koruna = sum(koruna_count),
    n_exchange = sum(exchange_count),
    n_stockshare = sum(stockshare_count),
    n_credit_loan = sum(credit_loan_count),
    n_foreclosure = sum(foreclosure_count),
    n_poverty = sum(poverty_count)
  ) %>%
  ungroup() %>%
  # Add a row with the total counts for each economy term across all years
  bind_rows(summarise(
    ., across(where(is.numeric), sum),
    across(where(is.character), ~"Total")
  ))

# This code chunk uses the kable function from the knitr package to create a table of the term counts per year in markdown format.
# The resulting table is printed below the code chunk in the R Markdown document.
knitr::kable(economy_term_counts_per_year, format = "pipe")

```

## Visualize all of the terms and their frequencies
```{r}
# This code reads in a data frame of economy term counts per year, filters out the "Total" row, 
# pivots the data frame to a longer format, removes the "n_" prefix from the economy term column, 
# and creates a grouped bar chart using ggplot2 to display the lemma occurrences per year. 
# The x-axis represents the year, the y-axis represents the count of lemma occurrences, 
# and the fill represents the lemma economy term. The chart is displayed with a minimal theme, 
# a continuous y-axis with breaks and labels every 500 counts, and a color palette from 
# the RColorBrewer package. The x-axis label is hidden, the y-axis label is "Lemma occurrences per year", 
# and the fill legend is blank.

economy_term_counts_per_year %>%
  filter(year != "Total") %>% 
  pivot_longer(where(is.numeric), values_to = "count", names_to = "economy_term") %>% 
  mutate(year = as.factor(year),
         economy_term = str_remove(economy_term, "^n_")) %>% 
  ggplot(aes(x = year, y = count, fill = economy_term)) +
  geom_col() +
   theme_minimal() +
    scale_y_continuous(
      breaks = seq(0, 50000, 2000),
      labels = seq(0, 50000, 2000)
    ) +
    scale_fill_brewer(palette = "Set1")+ 
    labs(x = element_blank(),
             y = "Lemma occurences per year",
             fill = "")
ggsave("economy_term_counts_plot.png", width = 18, height = 12, dpi = 600)
```
