---
title: "Frequency Analysis of Lemmata"
author: "Your Name"
date: "2021-10-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# This R Markdown file includes six sections: 
# The YAML header, which specifies the document title, author, date, and output format.
# The "setup" chunk, which sets some global options for the document.
# The "load-libraries" chunk, which loads the necessary R packages.
# The "load-stopwords" chunk, which loads the stopword lists from text files.
# The "load-data" chunk, which loads the data from the corpus files and combines them into a single data frame.
# The "clean-data" chunk, which removes stopwords, numbers, and interpunction from the lemma variable in the data frame.
# The "count-lemmata" chunk, which counts the number of occurrences of each lemma in each corpus, samples the top 1000 and 5000 lemmata, and writes the results to CSV files.
# The "plot-lemmata" chunk, which loads the top 1000 lemmata from the CSV file and creates a bar chart of the top 20 lemmata, with formatted y-axis labels

library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(scales)
```
##Now we will load the data and clean it
# Load the stopword lists
```{r setup, include=FALSE}
stop_words_cs <- read_lines("StopwordsCS.txt")
stop_words_czech_television <- read_lines("StopwordsCTV.txt")

# Create a list of corpus names
```{r setup, include=FALSE} 

corpus_names <- c()
for (year in 2012:2022) {
    if (year == 2022) {
        corpus_names <- c(corpus_names, paste0("Corpus", year - 2000, "_", sprintf("%02d", 1:4)))
    } else {
        corpus_names <- c(corpus_names, paste0("Corpus", year - 2000, "_", sprintf("%02d", 1:12)))
    }
}
```
# Create a list of file paths
```{r setup, include=FALSE} 
file_paths <- list.files("~/Rprojekt/Climate/2.data_transformations/data/udpipe_processed/", pattern = "udpipe_regex_.*\\.rds", full.names = TRUE)

# Loop through the file paths and create the corpora
for (i in seq_along(file_paths)) {
    corpus_name <- str_extract(file_paths[i], "Corpus\\d{2}_\\d{2}")
    if (corpus_name %in% corpus_names) {
        assign(corpus_name, readRDS(file_paths[i]) %>% select(doc_id, lemma))
    }
}

# Combine all corpora into a single data frame
corpus_df <- bind_rows(mget(corpus_names), .id = "corpus")

# Remove stopwords, numbers, and interpunction from the lemma variable
corpus_df_clean <- corpus_df %>%
    mutate(lemma = str_remove_all(lemma, "\\d+")) %>%
    mutate(lemma = str_remove_all(lemma, "[[:punct:]]+")) %>%
    filter(!lemma %in% stop_words_cs) %>%
    filter(!lemma %in% stop_words_czech_television)

```
# Count the number of occurrences of each lemma in each corpus, sample the top 1000 and 5000 lemmata, and write the results to CSV files

```{r setup, include=FALSE} 
lemma_counts <- corpus_df_clean %>%
    group_by(corpus, lemma) %>%
    summarise(count = n()) %>%
    pivot_wider(names_from = "corpus", values_from = "count", values_fill = 0)

lemma_counts <- lemma_counts %>% filter(!lemma == "NA")

# Sample the top 1000 lemmata
top_lemmata <- lemma_counts %>%
    group_by(lemma) %>%
    summarise(total_count = sum(across(everything()))) %>%
    top_n(1000, total_count) %>%
    inner_join(lemma_counts, by = "lemma")

# Write to CSV
write.csv(top_lemmata, "top_1000_lemmata.csv", row.names = FALSE)

# Sample the top 5000 lemmata
top_5000_lemmata <- lemma_counts %>%
    group_by(lemma) %>%
    summarise(total_count = sum(across(everything()))) %>%
    top_n(5000, total_count) %>%
    inner_join(lemma_counts, by = "lemma")

# Write to CSV
write.csv(top_5000_lemmata, "top_5000_lemmata.csv", row.names = FALSE)
```
# We will also include specific lemmata for climate issues.
```{r klima_klimaticky}
#This one is for the two lemmata for climate:
climate_noun_adj_lemmafreq <- lemma_counts %>%
    group_by(lemma) %>%
    summarise(total_count = sum(across(everything()))) %>%
    filter(lemma %in% c("klima", "klimatický")) %>%
    inner_join(lemma_counts, by = "lemma")

# Write to CSV
write.csv(climate_lemmafreq, "climate_noun_adj_lemma_frequencies.csv", row.names = FALSE)
```


```{r other climate lemmata}
#Specific lemmata that Andrea Culková asked for:
climate_lemmafreq <- lemma_counts %>%
    group_by(lemma) %>%
    summarise(total_count = sum(across(everything()))) %>%
    filter(lemma %in% c("ekologie", "ekolog", "environment", "environmentální", "znečištění", "nízkoemisní", "uhlík", "uhlíkový", "co2", "fosilní", "dekarbonizace", "vedra", "vedro", "oteplování", "povodeň", "tornádo")) %>%
    inner_join(lemma_counts, by = "lemma")

# Write to CSV
write.csv(climate_lemmafreq, "climate_lemma_frequencies.csv", row.names = FALSE)
```
# Now for the visualisations. We will create a bar chart of the top 20 lemmata, with formatted y-axis labels. Lemma 1 is a space character, so this is left out.

```{r top-lemmata}
top_lemmata_sorted <- top_lemmata[order(-top_lemmata$total_count),]
top_lemmata_sorted <- top_lemmata_sorted %>% filter(!lemma == "NA")
#MUST REMOVE NA!
ggplot(top_lemmata_sorted[2:21, ], aes(x = reorder(lemma, -total_count), y = total_count, fill = lemma)) +
    geom_bar(stat = "identity") +
    xlab("Lemma") +
    ylab("Total Count") +
    ggtitle("Top 20 Lemmata") +
    scale_y_continuous(labels = function(x) {
        ifelse(x >= 1e6, paste0(x / 1e6, " 000 000"),
                     ifelse(x >= 1e3, paste0(x / 1e3, " 000"), as.character(x)))
    })
ggsave("Top20Lemmata.png", width = 15, height = 10, dpi = 600)
```
    
    #And two other bar charts for the climate lemmata
```{r graph klima klimaticky}
ggplot(climate_noun_adj_lemmafreq, aes(x = reorder(lemma, -total_count), y = total_count, fill = lemma)) +
  geom_bar(stat = "identity", width = 0.7) +
  xlab("Lemma") +
  ylab("Total Count") +
  ggtitle("Frequency of 'klima' (noun) and 'klimatický' (adj.)") +
  scale_y_continuous(labels = function(x) {
    ifelse(x >= 1e6, paste0(x / 1e6, " 000 000"),
           ifelse(x >= 1e3, paste0(x / 1e3, " 000"), as.character(x)))
  }) +
  scale_fill_manual(name = "Climate noun & adjective", values = c("#7EFF0A", "#00770A"))

ggsave ("ClimateLemmata.png", width = 18, height = 12, dpi = 600)
```

    ```{r other climate_lemmata}
english_labels <- c("Flood(ing)", "Tornado", "Pollution", "Ecologist", "Heat(wave)", "Ecology", "Warming", "Fossil (Adj.)", "Carbon (Adj.)", "Low-emission (Adj.)", "Carbon (Noun)", "Environmental", "De-carbonization", "Environment")

# Define the order of the lemmata
lemma_order <- c("povodeň", "tornádo", "znečištění", "ekolog", "vedro", "ekologie", "oteplování", "fosilní", "uhlíkový", "nízkoemisní", "uhlík", "environmentální", "dekarbonizace", "environment")


# Define a custom color palette
custom_palette <- c("#1f77b4", "#aec7e8", "#000000", "#66bd63", "#d73027", "#1a9850", "#f46d43", "#969696", "#525252", "#e6f5d0", "#737373", "#d9ef8b", "#a6d96a", "#66c2a5")
names(custom_palette) = c("povodeň", "tornádo", "znečištění", "ekolog", "vedro", "ekologie", "oteplování", "fosilní", "uhlíkový", "nízkoemisní", "uhlík", "environmentální", "dekarbonizace", "environment")
english_translations <- c("povodeň"="Flood(ing)", "tornádo"="Tornado", "znečištění"="Pollution", "ekolog"="Ecologist", "vedro"="Heat(wave)", "ekologie"="Ecology", "oteplování"="Warming", "fosilní"="Fossil (Adj.)", "uhlíkový"="Carbon (Adj.)", "nízkoemisní"="Low-emission (Adj.)", "uhlík"="Carbon (Noun)", "environmentální"="Environmental", "dekarbonizace"="De-carbonization", "environment"="Environment")

climate_lemmafreq_sorted$lemma_factor <- factor(climate_lemmafreq_sorted$lemma, levels = unique(climate_lemmafreq_sorted$lemma))

ggplot(climate_lemmafreq_sorted, aes(x = reorder(lemma_factor, -total_count), y = total_count, fill = lemma_factor)) +
  geom_bar(stat = "identity") +
  xlab("Lemma") +
  ylab("Total Count") +
  ggtitle("Frequency of specific climate lemmata") +
  scale_y_continuous(labels = function(x) {
    ifelse(x >= 1e6, paste0(x / 1e6, " 000 000"),
           ifelse(x >= 1e3, paste0(x / 1e3, " 000"), as.character(x)))
  }) +
  scale_fill_manual(name = "English Labels", values = custom_palette, 
                    labels = climate_lemmafreq_sorted$english_labels[match(unique(climate_lemmafreq_sorted$lemma), climate_lemmafreq_sorted$lemma)])

ggsave("ClimateRelatedLemmata.png", width = 21, height = 14, dpi = 600)

```
