---
title: "frequencies2024"
author: "Irene Elmerot / Ondra Pekáček"
date: "2024-02-12"
output: html_document
---

---
title: "Word frequencies"
---
# This new version of a R Markdown file contains code for analyzing the frequency of selected lemmatized terms in Czech Television news articles from January 2012 to April 2022. 
# The code loads necessary packages, processes the data, and creates visualizations of the frequency of the selected terms over time. 
# The code also calculates the monthly proportion of overall documents containing at least one of the climate-terms. 
# The resulting visualizations are saved as PNG files in the "visuals" folder.

#Load necessary packages
```{r include=FALSE}
# Package names
packages <-
  c(
    "dplyr",
    "stringr",
    "purrr",
    "tidyr",
    "tidytext",
    "quanteda",
    "jsonlite",
    "ggplot2",
    "data.table",
    "plotly",
    "forcats",
    "ggwordcloud",
    "ggpubr"
  )

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, library, character.only = TRUE))
```

```{r The terms, if needed.}

# To include bi- or trigrams, the line summarize(text = str_flatten(climate_string_matches, collapse = " ")) is needed in the next chunk.
terms <- c("změna klima", "klima se měnit", "měnit se klima", "globální oteplování", "oteplování planeta", "klimatický změna", "klimatický podmínka", "klimatický dopad", "klimatický důsledek", "klimatický model", "klimatický opatření", "klimatický krize", "klimatický kolaps", "klimatický katastrofa", "klimaticky neutrální", "klimatický neutralita", "klimatický plán", "klimatický rozvrat", "klimatický zákon", "klimatický závazek", "klimatický žaloba", "klimatický vzdělávání")

```


```{r Creating the monthly numbers}

process_df <- function(df) {
  df %>% 
    transmute(article_id, climate_string_matches = tolower(climate_string_matches)) %>% 
    group_by(article_id) %>%
    summarize(text = str_flatten(climate_string_matches, collapse = " ")) %>%
    filter(map_lgl(text, ~ any(str_detect(.x, terms))))
}

climate_terms <- process_df(climate_corpus_v4)

climate_corpus_v4 <- climate_corpus_v4 %>% mutate(doc_id = article_id)

monthly_climate_counts <- doc_id_by_date %>% 
  inner_join(climate_corpus_v4, by = c("doc_id" = "article_id")) %>% 
  select(doc_id, date)

library(dplyr)
library(ggplot2)
library(lubridate)

# Extract the year and month from the date column and count the number of documents for each year and month
monthly_climate_counts <- monthly_climate_counts %>%
  mutate(year_month = floor_date(date, "month")) %>%
  count(year_month)

# Create the plot
ggplot(monthly_climate_counts, aes(x = year_month, y = n)) +
  geom_line(col = "darkgreen") +
   ylab("Document counts") +
  xlab(element_blank()) +
  labs(title = "A decade of Czech Television news",
       subtitle = "Total monthly counts of documents containing at least one of the lemmatized terms, 2012-2022",
       caption = "Source: Newton Media Archive, January 2012 to April 2022. N=2914.") +
  theme_classic2() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text = element_text(size = 7),
        plot.background = element_rect(fill = "grey90"),
        plot.title = element_text(face = "bold", size = 10),
        plot.subtitle = element_text(face = "italic", size = 8),
        plot.caption = element_text(size = 5),
        axis.title.y = element_text(size = 8),
        legend.title = element_blank(),
        legend.key = element_rect(fill = NA, linewidth = 8), #The `size` argument of `element_rect()` is deprecated as of ggplot2 3.4.0.Please use the `linewidth` argument instead
        legend.text = element_text(size = 9),
        legend.background = element_rect(fill = NA),
        legend.position = c(0.85, 0.9),
        plot.margin = margin(7,30,3,5, "pt")) +
  scale_x_date(date_breaks = "6 months", date_labels = "%m-%y") +
  coord_cartesian(xlim = c(start_date, end_date + 120), expand = FALSE) 

ggsave("~/Rprojekt/Climate/3.data_exploration/visuals/climate_terms_monthly_frequencies.png", device = "png",
       width = 1920, height = 1080, units = "px")
```

```{r Compare all with the climate mentions}
monthly_climate_counts <- monthly_climate_counts %>% mutate(monthly_count_climate = n)

all_articles_per_month <- doc_id_by_date %>%
    mutate(year_month = floor_date(date, "month")) %>%
    count(year_month, name = "monthly_count_all")

# Combine both datasets and save
combined_counts_per_month <- inner_join(all_articles_per_month, monthly_climate_counts, by = c("year_month" = "year_month")) %>%
      mutate(climate_prop = monthly_count_climate/monthly_count_all) %>% select(!n)

saveRDS(combined_counts_per_month, "~/Rprojekt/Climate/3.data_exploration/data/combined_counts_per_month.rds")
```
```{r}
# Convenience function to give ceiling to decimals for graph size purposes
ceiling_dec <- function(x, level=1) round(x + 5*10^(-level-1), level)
```

```{r Plot the comparison}
combined_counts_per_month %>% 
  ggplot(aes(x = year_month, y = climate_prop)) +
  geom_line(color = "#13a576") +
  geom_point(color = "#13a576", size = 0.3) +
  ylab(element_blank()) +
  xlab(element_blank()) +
  labs(title = "Decade of Czech Television news",
       subtitle = "Monthly proportion of overall documents containing at least one of the climate-terms",
       caption = "Source: Newton Media Archive, January 2012-April 2022. N=531,592. Selected lemmata are in climate_corpus_v4") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text = element_text(size = 7),
        plot.background = element_rect(fill = "grey90"),
        plot.title = element_text(face = "bold", size = 10),
        plot.subtitle = element_text(face = "italic", size = 8),
        plot.caption = element_text(size = 4.5),
        axis.title.y = element_text(size = 8),
        legend.title = element_blank(),
        plot.margin = margin(7, 30, 3, 0, "pt")) +
  scale_x_date(date_breaks = "6 months", date_labels = "%m-%y") +
  coord_cartesian(xlim = c(start_date, end_date + 120), expand = FALSE) +
  scale_y_continuous(
    expand = c(0, 0),
    breaks = seq(0, 1, 0.01),
    labels = paste0(seq(0, 100, 1), "%"),
    limits = c(0, ceiling_dec(max(combined_counts_per_month$climate_prop), 2))
  )

ggsave("~/Rprojekt/Climate/3.data_exploration/visuals/climate_lemma_proportions.png", device = "png",
       width = 1920, height = 1080, units = "px")
```

```{r}
doc_id_list <- list()

library(purrr)
library(dplyr)

# Get the list of .rds files
files <- list.files("~/Rprojekt/Climate/2.data_transformations/data/udpipe_processed/", pattern = "\\.rds$", full.names = TRUE)

# Read each .rds file, extract the doc_id column, and combine them into a data frame
doc_id_df <- map_dfr(files, ~readRDS(.x) %>% select(doc_id))

# The doc_id_df had to be exported and the duplicates from the doc_id column removed (done in BBEdit)
write_csv2(doc_id_df, file = "Corpus_total_ids.csv", quote = "none")
#The unique ids stored as Corpus_total_unique_ids.csv and the number of rows are the number of shows.

All_shows <-  531593/1000 #Count of all unique document IDs, divided by thousand to make the graph easier to read.
Climate_inclusive <- 2914/1000 #Count of all unique document IDs in climate_corpus_v4, divided by thousand to make the graph easier to read.
library(ggplot2)

# Create a data frame with the variables
documents_count <- data.frame(category = c("All_shows", "Climate_inclusive"),
                 value = c(531593/1000, 2914/1000))

# Create the bar plot
ggplot(documents_count, aes(x = category, y = value, fill = category)) +
  geom_bar(stat = "identity") +
  ylab("Nr of documents (thousands)") +
  xlab("All (100 %) vs climate change inclusive (0.5 %)") +
  labs(title = "Decade of Czech Television news",
       subtitle = "Documents (in thousands) containing selected lemma compared to all documents, 2012 - 2022",
       caption = "Source: Newton Media Archive, January 2012-April 2022, our climate_corpus_v4. All = 531,593, of which 2914 include climate terms.") +
  theme_bw() +
  theme(
    axis.text = element_text(size = 7),
    plot.background = element_rect(fill = "grey90"),
    plot.title = element_text(face = "bold", size = 10),
    plot.subtitle = element_text(face = "italic", size = 8),
    plot.caption = element_text(size = 4.5),
    axis.title.y = element_text(size = 8),
    legend.position = "none",
    plot.margin = margin(7, 30, 3, 0, "pt")
  ) +
  scale_fill_manual(values = c("#0F428A", "#009087", "#D53F3A"))



ggsave("~/Rprojekt/Climate/3.data_exploration/visuals/climate_lemma_counts.png", device = "png",
       width = 1920, height = 1080, units = "px")
  
```

