---
title: "Synchronize local with remote Google Drive backup repository"
params:
  remote_path: "climate_change_datasets/datasets/"
---
This R Markdown file contains code to synchronize local folders with a remote Google Drive backup repository. 
The code uses the `google_drive_sync.R` script to authenticate the user using OAuth in the browser, and then synchronizes a list of selected folders between the local machine and the remote repository. 
The list of folders to synchronize is defined in the `list_of_folders_to_synchronize` object, which contains the local and remote paths for each folder. 
The `synchronize_selected_folders()` function is then called with this list as an argument to perform the synchronization. 

The `params` object at the top of the file defines a single parameter `remote_path`, which specifies the path to the remote folder in the Google Drive repository where the files should be synchronized. 
This parameter can be set by the user when running the code.

# Setup
```{r}
source("google_drive_sync.R")

# Authenticate using OAuth in the browser
drive_auth()

# Current user info
drive_user()

# Open web browser
drive_browse(params$remote_path)

```


Example usage
```{r}
list_of_folders_to_synchronize <-
  list(
    media_articles_full = c(
      local_folder_path = "../1.data_extraction/data/full_articles/",
      remote_folder_path = paste0(params$remote_path, "1.extracted_data/full_articles/")
    ),
    media_articles_annotations = c(
      local_folder_path = "../1.data_extraction/data/annotated_articles/",
      remote_folder_path = paste0(params$remote_path, "1.extracted_data/annotated_articles/")
    ),
    media_articles_processed_regex = c(
      local_folder_path = "../2.data_transformations/data/regex_processed/",
      remote_folder_path = paste0(params$remote_path, "2.transformed_data/regex_processed/")
    ),
    media_articles_processed_udpipe = c(
      local_folder_path = "../2.data_transformations/data/udpipe_processed/",
      remote_folder_path = paste0(params$remote_path, "2.transformed_data/udpipe_processed/")
    ),
    media_articles_processed_subcorpus = c(
      local_folder_path = "../2.data_transformations/data/climate_sub_corpus/",
      remote_folder_path = paste0(params$remote_path, "2.transformed_data/climate_sub_corpus/")
    ),
    media_articles_analyzed_ner = c(
      local_folder_path = "../4.data_analysis/named_entity_recognition/data/",
      remote_folder_path = paste0(params$remote_path, "4.analyzed_data/named_entity_recognition/")
    ),
    media_articles_analyzed_sentiment = c(
      local_folder_path = "../4.data_analysis/sentiment_analysis/data/",
      remote_folder_path = paste0(params$remote_path, "4.analyzed_data/sentiment_analysis/")
    ),
     media_articles_analyzed_topic = c(
      local_folder_path = "../4.data_analysis/topic_modeling/data/",
      remote_folder_path = paste0(params$remote_path, "4.analyzed_data/topic_modeling/")
    ),
     media_articles_analyzed_topic_models_lda = c(
      local_folder_path = "../4.data_analysis/topic_modeling/data/models/lda",
      remote_folder_path = paste0(params$remote_path, "4.analyzed_data/topic_modeling/models/lda")
    ),
     media_articles_analyzed_topic_models_nmf = c(
      local_folder_path = "../4.data_analysis/topic_modeling/data/models/nmf",
      remote_folder_path = paste0(params$remote_path, "4.analyzed_data/topic_modeling/models/nmf")
    )
  )

synchronize_selected_folders(list_of_folders_to_synchronize)

```

